---
title: "Ferrum Manuscript Statistics"
output: html_document
date: "2025-11-14"
---
 #Load packages
```{r}
library(dataRetrieval)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(emmeans)
library(lme4)
library(ggeasy)
library(lattice)
library(rstatix)
library(ggsignif)
library(outliers)
library(ggpmisc)
library(ggpubr)
library(r2symbols)
library(stargazer)
library(ggbreak)
library(writexl)
library(ggrepel)
library(stringr)
library(ggthemes)
library(reshape)
library(RColorBrewer)
library(readxl)
library(dplyr)
library(gridExtra)
library(scales)
library(dbscan)
library(mclust)
library(devtools)
library(factoextra)
library(corrplot)
library(ggbiplot)
library(ggcorrplot)
library(GGally)
library(Hmisc)
library(ggbiplot)
library(ggfortify)
library(emmeans)
library(lme4)
library(lmerTest)
library(multcomp)
library(readxl)
library(learnr)
library(scatterplot3d)
library(vegan)
library(ggtern)
library(plotrix)
library(smwrBase)
library(smwrData)
library(smwrGraphs)
library(colorBlindness)
library(flextable)
library(officer)
library(dunn.test)
library(multcompView)
library(patchwork)
library(ggpattern)
```

#Load Data
```{r}
Alaska_DataRelease_2022_2023_Ferrum_V10 <- read_excel("C:/Users/tevinger/Box/Poulin Lab ETOX/Project Folders/Alaska BITE_HEAT (Taylor Evinger)/Evinger Manuscripts/Ferrum Manuscript/Datasheets/Ferrum Manuscript Datasheets/Alaska_DataRelease_2022_2023_Ferrum_V10.xlsx")

Alaska_DataRelease_2022_2023_Ferrum_V11_molar <- read_excel("C:/Users/tevinger/Box/Poulin Lab ETOX/Project Folders/Alaska BITE_HEAT (Taylor Evinger)/Evinger Manuscripts/Ferrum Manuscript/Datasheets/Ferrum Manuscript Datasheets/Alaska_DataRelease_2022_2023_Ferrum_V11_molar.xlsx")
```

# Statistics Codes

#Summary Stats Set Up
## Create specific dataframes
### Site Classification Group Data Only
```{r}
# filter out samples without a New_Group - use for summary stats of the groups
Grouped_Ferrum_2022_2023_data <- 
  Alaska_DataRelease_2022_2023_Ferrum_V10 %>%
  #Alaska_DataRelease_2022_2023_Ferrum_V11_molar %>%
  dplyr::select(ParkID,
                Watershed,
                #Subcatchment,
                New_Groups,
                New_Grouping,
                Field_label,
                Latitude,
                Longitude,
                sample_collection_year,
                sample_collection_month,
                sample_collection_season,
                Watershed_Area,
                Relative_MS_Watershed_Area,
                SamplingEventID,
                VisualDescription,
                Site_Classification,
                Hyd_Classification,
                Distance_km,
                Elevation_ft,
                pH,
                Temp_deg_celsius,
                Diss_oxy_mg_per_l,
                Diss_oxy_percent_sat,
                Spec_Cond_microS_per_cm,
                DIC_mgC_per_l,
                HCO3_mg_per_l,
                Alk_mgCaCO3_per_l,
                DOC_mgC_per_l,
                f_Cl_mg_per_l,
                f_NO3_mgN_per_l,
                f_SO4_mg_per_l,
                f_Ca_mg_l,
                f_Mg_mg_l,
                f_Na_mg_l,
                f_K_mg_l,
                f_SiO2_mg_per_l,
                f_Pb_mcg_per_l, p_Pb_mcg_per_l,
                f_Ag_mcg_per_l, p_Ag_mcg_per_l,
f_Al_mcg_per_l, p_Al_mcg_per_l,
f_As_mcg_per_l, p_As_mcg_per_l,
f_Ba_mcg_per_l, p_Ba_mcg_per_l,
#f_Be_mcg_per_l, p_Be_mcg_per_l,
f_Cd_mcg_per_l, p_Cd_mcg_per_l,
f_Ce_mcg_per_l, p_Ce_mcg_per_l,
f_Co_mcg_per_l, p_Co_mcg_per_l,
f_Cr_mcg_per_l, p_Cr_mcg_per_l,
f_Cu_mcg_per_l, p_Cu_mcg_per_l,
f_Dy_mcg_per_l, p_Dy_mcg_per_l,
f_Fe_mcg_per_l, p_Fe_mcg_per_l,
f_La_mcg_per_l, p_La_mcg_per_l,
f_Mn_mcg_per_l, p_Mn_mcg_per_l,
f_Nd_mcg_per_l, p_Nd_mcg_per_l,
f_Ni_mcg_per_l, p_Ni_mcg_per_l,
f_Pr_mcg_per_l, p_Pr_mcg_per_l,
f_Se_mcg_per_l, p_Se_mcg_per_l,
#f_Th_mcg_per_l, p_Th_mcg_per_l,
f_Tl_mcg_per_l, p_Tl_mcg_per_l,
f_U_mcg_per_l, p_U_mcg_per_l,
f_V_mcg_per_l, p_V_mcg_per_l,
f_Y_mcg_per_l, p_Y_mcg_per_l,
f_Zn_mcg_per_l, p_Zn_mcg_per_l,
f_REE, p_REE,
f_sum_all_metals, 
u_sum_all_metals,
p_sum_all_metals) %>%
  filter(!(is.na(New_Grouping)))

Grouped_Ferrum_2022_2023_data <- Grouped_Ferrum_2022_2023_data %>%
  mutate(across(19:86, as.numeric))

view(Grouped_Ferrum_2022_2023_data)
```

# Summary Statistics
## set elements and columns for summary stats loop
```{r}
#summary stats
library(dplyr)

# Define the elements and corresponding column names
#elements <- c("pH", "SpC", "SO4", "DIC", "Alk", "DOC", "Ca", "Mg", "Na", "K", "Cl", "NO3", "DO")
#columns <- c("pH", "Spec_Cond_microS_per_cm", "f_SO4_mg_per_l", "DIC_mgC_per_l", "Alk_mgCaCO3_per_l", "DOC_mgC_per_l", "f_Ca_mg_l", "f_Mg_mg_l", "f_Na_mg_l", "f_K_mg_l", "f_Cl_mg_per_l", "f_NO3_mg_per_l", "Diss_oxy_mg_per_l")

elements <- c(
  "pH", "Temp", "Diss_oxy", "Diss_oxy_percent_sat", "Spec_Cond", "DIC", "HCO3",
  "Alk", "DOC", "f_Cl", "f_NO3", "f_SO4", "f_Ca", "f_Mg", "f_Na", "f_K",
  "f_SiO2",
 "f_Pb", "p_Pb", 
 "f_Ag", "p_Ag",
"f_Al", "p_Al",
"f_As", "p_As",
"f_Ba", "p_Ba",
#"f_Be", "p_Be",
"f_Cd", "p_Cd",
"f_Ce", "p_Ce",
"f_Co", "p_Co",
"f_Cr", "p_Cr",
"f_Cu", "p_Cu",
"f_Dy", "p_Dy",
"f_Fe", "p_Fe",
"f_La", "p_La",
"f_Mn", "p_Mn",
"f_Nd", "p_Nd",
"f_Ni", "p_Ni",
"f_Pr", "p_Pr",
"f_Se", "p_Se",
#"f_Th", "p_Th",
"f_Tl", "p_Tl",
"f_U", "p_U",
"f_V", "p_V",
"f_Y", "p_Y",
"f_Zn", "p_Zn",
"f_REE", "p_REE",
"f_sum_of_metals", "p_sum_of_metals"
)

columns <- c(
  "pH", "Temp_deg_celsius", "Diss_oxy_mg_per_l", "Diss_oxy_percent_sat", "Spec_Cond_microS_per_cm", 
  "DIC_mgC_per_l", "HCO3_mg_per_l",  "Alk_mgCaCO3_per_l", "DOC_mgC_per_l", "f_Cl_mg_per_l", "f_NO3_mgN_per_l", 
  "f_SO4_mg_per_l", "f_Ca_mg_l", "f_Mg_mg_l", "f_Na_mg_l", "f_K_mg_l",
  "f_SiO2_mg_per_l",
  "f_Pb_mcg_per_l", "p_Pb_mcg_per_l", 
  "f_Ag_mcg_per_l", "p_Ag_mcg_per_l",
"f_Al_mcg_per_l", "p_Al_mcg_per_l",
"f_As_mcg_per_l", "p_As_mcg_per_l",
"f_Ba_mcg_per_l", "p_Ba_mcg_per_l",
#"f_Be_mcg_per_l", "p_Be_mcg_per_l",
#"f_Br_mcg_per_l", "p_Br_mcg_per_l",
"f_Cd_mcg_per_l", "p_Cd_mcg_per_l",
"f_Ce_mcg_per_l", "p_Ce_mcg_per_l",
"f_Co_mcg_per_l", "p_Co_mcg_per_l",
"f_Cr_mcg_per_l", "p_Cr_mcg_per_l",
"f_Cu_mcg_per_l", "p_Cu_mcg_per_l",
"f_Dy_mcg_per_l", "p_Dy_mcg_per_l",
"f_Fe_mcg_per_l", "p_Fe_mcg_per_l",
"f_La_mcg_per_l", "p_La_mcg_per_l",
"f_Mn_mcg_per_l", "p_Mn_mcg_per_l",
"f_Nd_mcg_per_l", "p_Nd_mcg_per_l",
"f_Ni_mcg_per_l", "p_Ni_mcg_per_l",
"f_Pr_mcg_per_l", "p_Pr_mcg_per_l",
"f_Se_mcg_per_l", "p_Se_mcg_per_l",
#"f_Th_mcg_per_l", "p_Th_mcg_per_l",
"f_Tl_mcg_per_l", "p_Tl_mcg_per_l",
"f_U_mcg_per_l", "p_U_mcg_per_l",
"f_V_mcg_per_l", "p_V_mcg_per_l",
"f_Y_mcg_per_l", "p_Y_mcg_per_l",
"f_Zn_mcg_per_l", "p_Zn_mcg_per_l",
"f_REE", "p_REE",
"f_sum_all_metals", "p_sum_all_metals"
)
```

## rstatix summary stats
```{r}
# Trying to make a loop for summary stats using rstatix instead of base R
library(dplyr)
library(rstatix)


# Initialize an empty list to store results
summary_stats_list <- list()
quartiles_list <- list()

# Loop through elements and columns
for (i in seq_along(elements)) {
  element <- elements[i]
  column <- columns[i]
  
  # Perform summary operation with a default empty tibble if an error occurs
  summary <- 
    #stats_working_Ferrum_data %>% # use for all data, watershed, year and month summary stats
    Grouped_Ferrum_2022_2023_data %>% # to group by New Grouping (seeps included)
    
    #dplyr::select(9:70) %>% # all data
    #dplyr::select(2,9:70) %>% # Watershed
    #dplyr::select(4,9:70) %>% # Year
    #dplyr::select(5,9:70) %>% # Month
    #dplyr::select(6,9:70) %>% # Season
    dplyr::select(2,4,19:86) %>% # New Grouping
    filter(!is.na(.data[[column]])) %>%
     group_by(New_Grouping
              #, Watershed
              ) %>% #This is where you can change what variable you want to group the data by
    #group_by(New_Grouping) %>% #This is where you can change what variable you want to group the data by
      get_summary_stats(vars = column, type = "common")

  
  # Calculate Q1 and Q3
  q1_q3 <- 
    #stats_working_Ferrum_data %>% # use for all data, watershed, year and month summary stats
    Grouped_Ferrum_2022_2023_data %>% # to group by New Grouping (seeps included)
    #dplyr::select(9:70) %>% # all data
    #dplyr::select(2,9:70) %>% # Watershed
    #dplyr::select(4,9:70) %>% # Year
    #dplyr::select(5,9:70) %>% # Month
    #dplyr::select(6,9:70) %>% # Season
    dplyr::select(2,4,19:86) %>% # New Grouping
    filter(!is.na(.data[[column]])) %>%
    group_by(New_Grouping
             #, Watershed
             ) %>% #This is where you can change what variable you want to group the data by
    #group_by(New_Grouping) %>% #This is where you can change what variable you want to group the data by
      summarise(
        variable = column,
        Q1 = round(quantile(.data[[column]], probs = 0.25, na.rm = TRUE), 2),
        Q3 = round(quantile(.data[[column]], probs = 0.75, na.rm = TRUE), 2),
        .groups = "drop"
      )
  
  # Store result
  summary_stats_list[[i]] <- summary
  quartiles_list[[i]] <- q1_q3
}

# Combine into one dataframe
summary_df_2 <- do.call(rbind, summary_stats_list)
quartiles_df <- do.call(rbind, quartiles_list)

quartiles_df <- quartiles_df %>%
  dplyr::rename(Variable = variable) %>%
  mutate(joining = paste(
    #Watershed, 
    New_Grouping, 
    Variable, sep = "_"))

summary_df_2 <- summary_df_2 %>%
  dplyr::rename(Variable = variable) %>%
  mutate(joining = paste(
    #Watershed, 
    New_Grouping, 
    Variable, sep = "_"))

view(summary_df_2) # Grouped by New_Grouping
view(quartiles_df)

summary_combined <- summary_df_2 %>%
  left_join(quartiles_df, by = "joining") %>%
  dplyr::select(-joining, 
                #-Watershed.y, 
                -New_Grouping.y, -Variable.y) %>%
  dplyr::rename(New_Grouping = New_Grouping.x) %>%
  dplyr::rename(Variable = Variable.x)
  #dplyr::rename(Watershed = Watershed.x)

#arrange New_Grouping so it goes upstream to downstream
summary_combined <- summary_combined %>%
  # 1. Define the order of New_grouping
  mutate(New_Grouping = factor(New_Grouping,
                               levels = c("Upstream", "Unimpaired Tributary", "Seep", "Impaired Tributary", "Downstream MS"))) %>%
  # 2. Arrange rows first by Variable, then by the factor order of New_grouping
  arrange(Variable, New_Grouping)

view(summary_combined)
#view(summary_df) # grouped by New_Grouping only

summary_combined_watershed <- summary_combined

#write_xlsx(summary_combined, "C:/Users/tevinger/Downloads/Group and watershed summary stats_rstatix_updated Nov 13_2025.xlsx")
```

```{r}
stats_Upstream <- summary_combined %>%
  dplyr::select(New_Grouping, Variable, n, median, sd) %>%   # keep only what you need
 filter(New_Grouping == "Upstream")

stats_UnimpairedTrib <- summary_combined %>%
  dplyr::select(New_Grouping, Variable, n, median, sd) %>%
  filter(New_Grouping == "Unimpaired Tributary")

stats_Seep <- summary_combined %>%
  dplyr::select(New_Grouping, Variable, n, median, sd) %>%
  filter(New_Grouping == "Seep")

stats_ImpairedTrib <- summary_combined %>%
  dplyr::select(New_Grouping, Variable, n, median, sd) %>%
  filter(New_Grouping == "Impaired Tributary")

stats_DownstreamMS <- summary_combined %>%
  dplyr::select(New_Grouping, Variable, n, median, sd) %>%
  filter(New_Grouping == "Downstream MS")

# Create a new workbook
wb <- createWorkbook()

# Add each dataframe as a sheet
addWorksheet(wb, "Upstream")
writeData(wb, "Upstream", stats_Upstream)

addWorksheet(wb, "UnimpairedTrib")
writeData(wb, "UnimpairedTrib", stats_UnimpairedTrib)

addWorksheet(wb, "Seep")
writeData(wb, "Seep", stats_Seep)

addWorksheet(wb, "ImpairedTrib")
writeData(wb, "ImpairedTrib", stats_ImpairedTrib)

addWorksheet(wb, "DownstreamMS")
writeData(wb, "DownstreamMS", stats_DownstreamMS)

# Save to Downloads (Windows/Mac compatible)
file_path <- file.path("C:/Users/tevinger/Downloads", "summary_stats_by_group.xlsx")
saveWorkbook(wb, file_path, overwrite = TRUE)

file_path
```

## subset Upstream, Seep, and Downstream MS groups
```{r}
# filter to keep only upstream and seep
summary_df_filtered <- summary_combined %>%
  filter(New_Grouping %in% c("Upstream", "Seep", "Downstream MS")) %>%
  mutate(New_Grouping = factor(New_Grouping, levels = c("Upstream", "Seep", "Downstream MS"))) %>%
  arrange(Variable, New_Grouping)

view(summary_df_filtered)

#write_xlsx(summary_df_filtered, "C:/Users/tevinger/Downloads/Upstream,seep,Downstream_filtered_rstatix.xlsx")


summary_df_filtered_wide <- summary_df_filtered %>%
  # select median (6) or mean (8)
  dplyr::select(1,2,8) %>%
  pivot_wider(
  id_cols = Variable,
  names_from = New_Grouping,
  values_from = mean
)

view(summary_df_filtered_wide)

#write_xlsx(summary_df_filtered_wide, "C:/Users/tevinger/Box/Poulin Lab ETOX/Project Folders/Alaska BITE_HEAT (Taylor Evinger)/Evinger Manuscripts/Ferrum Manuscript/Results and Discussion/Ferrum Manuscript Summary Stats/Upstream, seep, downstream_means.xlsx")
```

## subset Upstream and Downstream MS groups for each watershed
```{r}
# filter to keep only upstream and seep
summary_df_filtered <- summary_combined_watershed %>%
  filter(New_Grouping %in% c("Upstream", "Downstream MS")) %>%
  mutate(New_Grouping = factor(New_Grouping, levels = c("Upstream", "Downstream MS"))) %>%
  arrange(Watershed, Variable, New_Grouping)

view(summary_df_filtered)
```

```{r}
Upstream_particulate_metals <- summary_df_filtered %>%
  filter(New_Grouping %in% c("Upstream")) %>%
  dplyr::select(1,2,3,7) %>% # Watershed, New_Grouping, Variable, and median
  filter(str_starts(Variable, "p_")) %>%
  filter(median == 0.000) %>%
  mutate(
    median = case_when(
      # Group 1: Al, Mn, Ni, Cu, Zn → replace with sqrt(2)/2 * 0.1
      str_detect(Variable, "(Mn|Ni|Zn)") & median == 0.000 ~ (sqrt(2)/2) * 0.1,
      
      # Group 2: Fe → replace with sqrt(2)/2 * 1
      str_detect(Variable, "Fe") & median == 0.000 ~ (sqrt(2)/2) * 1,
      
      # Group 3: Be, V, Cr, Co, As, Se, Ag, Cd, Ba, Tl, Pb, Th, U, REE → sqrt(2)/2 * 0.01
      str_detect(Variable, "(Cr|Co|Se|Cd|Ba|U|sum)") & median == 0 ~ (sqrt(2)/2) * 0.01,
      
      # Group 4: REE → sqrt(2)/2 * 0.001
      str_detect(Variable, "(REE|Pr|Nd|Dy|La|Ce|Y)") & median == 0 ~ (sqrt(2)/2) * 0.001,
      
      # Else: keep value
      TRUE ~ median
    )
  )
```

```{r}
summary_df_filtered_median <- summary_df_filtered %>%
  dplyr::select(1,2,3,7)  # Watershed, New_Grouping, Variable, and median
```

```{r}
# Define key columns
key_cols <- c("Watershed", "New_Grouping", "Variable")

# Find all non-key columns that exist in BOTH data frames
common_value_cols <- intersect(
  setdiff(names(summary_df_filtered_median), key_cols),
  setdiff(names(Upstream_particulate_metals), key_cols)
)

common_value_cols
# just to eyeball: this is the set of columns that will be replaced

# 3. Join particulate values onto the summary df
joined <- summary_df_filtered_median %>%
  left_join(
    Upstream_particulate_metals,
    by = key_cols,
    suffix = c("", "_partic")  # summary: col, particulate: col_partic
  )

# 4. For each shared value column, replace with particulate where available
for (col in common_value_cols) {
  joined[[col]] <- dplyr::coalesce(
    joined[[paste0(col, "_partic")]],  # particulate version
    joined[[col]]                      # original version
  )
}

# 5. Drop the helper *_partic columns
summary_df_filtered_median <- joined %>%
  dplyr::select(-ends_with("_partic"))
```

```{r}
summary_df_filtered_wide <- summary_df_filtered_median %>%
  pivot_wider(
  id_cols = c(Watershed, Variable),
  names_from = New_Grouping,
  values_from = median
) %>%
  dplyr::rename(
    Upstream_median   = 3,
    Downstream_median = 4
  ) 
  

view(summary_df_filtered_wide)

#write_xlsx(summary_df_filtered_wide, "C:/Users/tevinger/Box/Poulin Lab ETOX/Project Folders/Alaska BITE_HEAT (Taylor Evinger)/Evinger Manuscripts/Ferrum Manuscript/Results and Discussion/Ferrum Manuscript Summary Stats/Upstream, seep, downstream_means.xlsx")
```

```{r}
Up_to_Downstream_differences <- summary_df_filtered_wide %>%
  mutate(downstream_diff_conc = Downstream_median - Upstream_median) %>%
  mutate(downstream_diff_fold_greater = (Downstream_median/Upstream_median)) %>%
  mutate(downstream_percent_increase = ((Downstream_median - Upstream_median)/Upstream_median) * 100)
```

```{r}
write_xlsx(Up_to_Downstream_differences, "C:/Users/tevinger/Downloads/Differences from Upstream to downstream.xlsx")
```

# Temporal Dataset Statistics
## Summary Stats
```{r}
#summary stats
library(dplyr)

# Define the elements and corresponding column names
elements <- c("pH", "SpC", "SO4", "DIC", "Alk", "DOC", "Ca", "Mg", "Na", "K", "Cl", "DO")

columns <- c("pH", "Spec_Cond_microS_per_cm", "f_SO4_mg_per_l", "DIC_mgC_per_l", "Alk_mgCaCO3_per_l", "DOC_mgC_per_l", "f_Ca_mg_l", "f_Mg_mg_l", "f_Na_mg_l", "f_K_mg_l", "f_Cl_mg_per_l", "Diss_oxy_mg_per_l")
```

## rstatix summary stats
```{r}

library(dplyr)
library(rstatix)

# Initialize an empty list to store results
summary_stats_list <- list()
quartiles_list <- list()

# Loop through elements and columns
for (i in seq_along(elements)) {
  element <- elements[i]
  column <- columns[i]
  
  # Perform summary operation with a default empty tibble if an error occurs
  summary <- 
    temporal_stats %>% # to group by New Grouping (seeps included)
    dplyr::select(9,12, columns) %>% # New Grouping
    filter(!is.na(.data[[column]])) %>%
     group_by(Temporal_status, sample_collection_season) %>% #This is where you can change what variable you want to group the data by
      get_summary_stats(vars = column, type = "common")

  
  # Calculate Q1 and Q3
  q1_q3 <- 
    temporal_stats %>% # to group by New Grouping (seeps included)
    dplyr::select(9,12, columns) %>% # New Grouping
    filter(!is.na(.data[[column]])) %>%
     group_by(Temporal_status, sample_collection_season) %>% #This is where you can change what variable you want to group the data by
      summarise(
        variable = column,
        Q1 = round(quantile(.data[[column]], probs = 0.25, na.rm = TRUE), 2),
        Q3 = round(quantile(.data[[column]], probs = 0.75, na.rm = TRUE), 2),
        .groups = "drop"
      )
  
  # Store result
  summary_stats_list[[i]] <- summary
  quartiles_list[[i]] <- q1_q3
}

# Combine into one dataframe
summary_df_2 <- do.call(rbind, summary_stats_list)
quartiles_df <- do.call(rbind, quartiles_list)

quartiles_df <- quartiles_df %>%
  dplyr::rename(Variable = variable) %>%
  mutate(joining = paste(
    Temporal_status, 
    sample_collection_season,
    Variable, 
    sep = "_"))

summary_df_2 <- summary_df_2 %>%
  dplyr::rename(Variable = variable) %>%
  mutate(joining = paste(
    Temporal_status, 
    sample_collection_season,
    Variable, 
    sep = "_"))

view(summary_df_2) # Grouped by New_Grouping
view(quartiles_df)

summary_combined <- summary_df_2 %>%
  left_join(quartiles_df, by = "joining") %>%
  dplyr::select(-joining,
                -Temporal_status.y, -Variable.y, -sample_collection_season.y) %>%
  dplyr::rename(Temporal_status = Temporal_status.x) %>%
  dplyr::rename(Variable = Variable.x) %>%
  dplyr::rename(sample_collection_season = sample_collection_season.x)

view(summary_combined)

summary_combined <- summary_combined %>%
  arrange(sample_collection_season, Variable)
#view(summary_df) # grouped by New_Grouping only

write_xlsx(summary_combined, "C:/Users/tevinger/Downloads/Aggie Temporal status_sample season summary stats.xlsx")
```

```{r}

```